### 2020-11-04 spectral radius of matrix new method/thesis/article110320.tex
1104 temp saved

\documentclass{article}

% \topmargin 0pt

% \topmargin=-1.5cm
% \oddsidemargin=0.7cm
% \textheight=23.5cm
% \textwidth=15cm
% \setlength{\baselineskip}{24pt}
% \renewcommand{\baselinestretch}{1.5} %行距

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
% \usepackage{amsfonts}
\usepackage{mathrsfs}
% \usepackage{tabularx,array}
% \usepackage{pgf,tikz}
% \usepackage{blkarray} %line 80-92 Q formula matrix index
%\usepackage{colortbl}

% \usetikzlibrary{arrows}
\setCJKmainfont{標楷體} %設定中文為系統上的字型，而英文不去更動，使用原TeX字型
% \XeTeXlinebreaklocale "zh"             %這兩行一定要加，中文才能自動換行
% \XeTeXlinebreakskip = 0pt plus1pt     %這兩行一定要加，中文才能自動換行
% \title{Combinatorial Identities from Lagrange's Interpolation Polynomial}
% \author{Student: Yen-Jung Huang  ~~~~~~~~~~~~~~~~~~~~~~~~~~Advisor: Chih-Wen Weng}
% \date{} %不要日期

\def\UrlFont{\rm}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[chapter]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{remark}[thm]{Remark}
\newtheorem{eg}[thm]{Example}
\newtheorem{conj}[thm]{Conjecture}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{ex}[thm]{Exercise}
\newtheorem{prob}[thm]{Problem}
\newtheorem{exam}[thm]{Example}
\newtheorem{nota}[thm]{Notation}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ques}[thm]{Question}
\newtheorem{pof}[thm]{Proof.}

\usepackage{comment}

%\renewcommand {\refname} {Bibliography}

\begin{document}

\section{Preliminaries}

    Let $\mathbb{R}$ and $\mathbb{C}$ denote the

field of real numbers and complex numbers respectively.
    \begin{defn}

Let $C$ be an $n \times n$ real nonnegative matrix, and $u \in \mathbb{R}^n$ be a

nonzero column vector. The scalar $\lambda \in \mathbb{C}$ is an $\it{eigenvalue}$

of C corresponding to the $\it{eigenvector}$ $u,$  if $Cu = \lambda u.$

%Also, $C$ has an left eigenvector $v^T$, when $v^TC=\lambda v^T$.
    \end{defn}

   \begin{defn}
        When $C$ is an $n \times n$ real matrix, the $\textit

{spectral radius} $ $\rho(C)$ of $C$ is defined by
        $$\rho(C):=\max\{~|

\lambda|~\ |~~\lambda\hbox{ is an eigenvalue of $C$}\},$$
        where $|

\lambda|$ is the magnitude of complex number $\lambda.$
    \end{defn}

\begin{thm} 
        %\bf{Perron–Frobenius theorem} \\
        If $C$ is

nonnegative square matrix, then the spectral radius $\rho(C)$ is an eigenvalue of

$C$ with a corresponding nonnogative right eigenvector and a corresponding

nonnegative left eigenvector.
    \end{thm}

    We are interested in spectral

radius of the following matrix associated with a simple graph.

    \begin{defn}

  Given an undirected graph G, the$\textit{ adjacency matrix}$ of G is the square

matrix $A = (a_{ij})$ indexed by vertices of G, and
    \[a_{ij} =\begin{cases}

 1, \text{if $i$ is adjacent to $j$}, \\
    0, \text{otherwise.}
    \end{cases}

   \]
    \end{defn}

    \begin{defn}
    Given an undirected graph G, the $

\textit{spectral radius}$  $\rho(G) $ of G is the spectral radius of the adjacency

matrix of G.
    \end{defn}

    We introduce a notation of submatrix, which is

taken from some columns and some rows of a matrix.
    \begin{defn}
        For a

matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column

indices of 
        $C$ respectively,  We use $C[\alpha|\beta]$ to denote the

submatrix of $C$ with size
         $ |\alpha| \times |\beta| $ that has entries

$c_{ij}$ for $i\in \alpha$ and $j\in\beta$,
    \end{defn}

We introduce two

matrices $P$ and $Q$ in the following theorem, where $P$ is a permutation matrix

which is multiplied to the left side, and $Q$ is sum of elementary matrix and

certain binary matrix. In which $P$ generalize row permutation on cases of $C$

matrix,and $Q$ is the transform from $C'$ to C', which is the first n-1 columns

and sum of certain columns. We aim to find C' such that $C'$ majors $C$,i.e. $C

\leq C'$

The following theorem is from [].\cite{content}

    \begin

{thm}\label{pre_thm}
        Let $C=(c_{ij})$, $C'=(c'_{ij})$, $P$ and $Q$ be  $n

\times n$ matrices.
        Assume that
        \begin{enumerate}

\item[(i)]    $PCQ\leq PC'Q$;\label{Q1}
            \item[(ii)]  there exist a

nonnegative column vector $u=(u_1, u_2, \ldots, u_n)^T$  and a scalar $\lambda'\in

\mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated

eigenvector $Qu$;
            \item[(iii)] there exist a nonnegative row vector

$v^T=(v_1, v_2, \ldots, v_n)$  and a scalar $\lambda\in \mathbb{R}$

such that $\lambda$ is an eigenvalue of $C$ with associated  left eigenvector

$v^TP$; and
            \item[(iv)] $v^TPQu>0.$
            \end{enumerate}

     Then $\lambda\leq \lambda'$.
                Moreover, $\lambda=\lambda'$

         if and only if
                \begin{equation}

\label{e3}
            (PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n

\hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
        \end{equation}
    \end

{thm}%

    % { Proof}
            Multiplying the nonnegative vector $u$ in

(ii) to the right of both terms of  (i),
        \begin{equation}\label{e1}

     PCQu\leq PC'Qu=\lambda'PQu.
        \end{equation}
        Multiplying the

nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption (iii) to the

left of all terms  in (\ref{e1}), we have
        \begin{equation}\label{e2}

      \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
        \end{equation}

       Now delete the positive term $v^TPQu$ by assumption (iv) to obtain $

\lambda\leq \lambda'$ and finish the proof of the first part.

    { Proof

(Continue)}
        Assume that $\lambda=\lambda'$, so the inequality in (\ref

{e2}) is an equality.  Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i

\not=0.$ Hence, $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$

with $u_j\not=0.$ 
         Conversely, (\ref{e3}) implies $$v^TPCQu=\sum_{i,j}

v_i(PCQ)_{ij}u_j=\sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so
            $

\lambda=\lambda'$ by (\ref{e2}).

\section{Our Method}
We use [n-1] as notation of the set of elements from one to n-1, which is {1,2,$...$,n}.
Throughout fix $k\in [n-1]$.Let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$ appearing in the position $k,n$ of $E_{kn}$. We will apply the previous theorem with $P=I$ and

\begin{equation}
Q=I+E_{kn}=\begin{pmatrix} \label{Q1}
1 &  & & &  & 0 \\
 & 1 &  &      &  &  \\
 &  & \ddots & &  & 1 \\
 &  &        & &  &  \\
  &  & & & 1 &  \\
0 &  & & &  & 1 \\
\end{pmatrix}.
\end{equation}

\begin{defn}[$k$-rooted vector]
~A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $k$-rooted}  if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n.$
\end{defn}
\bigskip

The following Lemma is immediate from the above definition.
\bigskip

\begin{lem}[vector rooted lemma]\label{v_rooted}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}
\item[(i)] $v'$ is $k$-rooted  if and only if  $u$ is nonnegative;
\item[(ii)] $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
\qed
\end{lem}

\begin{thm}[Our first result]\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}
\item[(i)]   $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item[(ii)] there exists a $k$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item[(iii)] there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item[(iv)] $v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}
    \item[(a)]    $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item[(b)]  $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
    \item[(c)]   $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$ 
\end{enumerate} \qed
\end{thm}
%

% {}

    \begin{proof}
        The proof is based on Theorem (\ref{pre_thm}) with $P = I$ and $Q = I + E_{kn} \text{ in } (\ref{Q1})$. 
        The assumption (i) $PCQ\leq PC'Q$ of Theorem (\ref{pre_thm}) holds by the condition (i) of this theorem. 
        Let $u = Q^{-1}v'$. Then u is nonnegative and $C'Qu = \lambda' Qu$ by the condition (ii) and 
        Lemma 2.2(i).  Hence the assumption (ii) of Theorem (\ref{pre_thm}) holds. The assumptions (iii) and (iv) 
        of Theorem (\ref{pre_thm}) clearly hold by conditions (iii),(iv) of this theorem since $P = I$ and $v'= Qu$  
        Hence $\lambda \leq \lambda' $ by the conclusion of Theorem (\ref{pre_thm}). Moreover $\lambda = \lambda'$ 
        if and only if (\ref{e3}) holds, and this is equivalent to conditions (a),(b) of this theorem. 
    \end{proof}

    We are interested in the matrices $C'$ that have $k$-rooted eigenvectors.
    Motivated by the condition (i) of theorem 2.3, we provide the following two definitions.

    \begin{defn}[(k,n)-sum]
        For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)$-{\it sum} vector of $C'$ is the vector of the sum of the $k$-th and  $n$-th columns of $C'$, where $k\leq n-1$.
    \end{defn}

    Note that the last column of $C'Q$ is the $(k, n)$-{\it sum} vector of $C'$

    \begin{defn}[k-rooted matrix]\label{m_rooted}
        A  matrix $C'=(c'_{ij})$ is called {\it $k$-rooted}  if its  columns and its $(k, n)$-sum vector are all $k$-rooted except the last column of $C'$.
    \end{defn}
%

%

    \begin{lem}\label{lma_m_rooted}
        Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(iii) hold.
            \begin{enumerate}
                \item[(i)]$C'$ is a $k$-rooted matrix, if and only if, $Q^{-1}C'Q$ is nonnegative.
                \item[(ii)]Assume that $C'$ is $k$-rooted and let $u$ be a nonnegative eigenvector of $Q^{-1}C'Q$ for $\rho(C')$. Then  $C'$ has a $k$-rooted eigenvector $v'=Qu$ for $\rho(C')$. 
                \item[(iii)] $\rho(C')$ = $\rho(Q^{-1}C'Q)$
            \end{enumerate}
    \end{lem}

%

    \begin{rem}

    $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
    1 &  & & &  & 0 \\
    & 1 &  &      &  &  \\
    &  & \ddots & &  & -1 \\
    &  &        & &  &  \\
    &  & & & 1 &  \\
    0 &  & & &  & 1 \\
    \end{pmatrix}.$$

    The matrix $C'Q$ is
    $$\begin{pmatrix}
    c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
    \vdots \\
    c'_{k-11}     & c'_{k-1 2}           & \cdots     & c'_{k-1 n-1} & c'_{k-1k}+c'_{k-1n} \\
    c'_{k1} & c'_{k2} &\cdots      & c'_{kn-1} & c'_{kk}+c'_{kn}\\
    c'_{k+11}     & c'_{k+12}           & \cdots     & c'_{k+1\ n-1} & c'_{k+1k}+c'_{k+1n} \\
    \vdots              & \vdots & \ddots              & \vdots & \vdots \\
    c'_{n1}             & c'_{n2} & \cdots             & c'_{n\ n-1} & c'_{nk}+c'_{nn} \\
    \end{pmatrix}.
    $$
    \end{rem}

\begin{comment}

{Remark}

	\begin{proof}
        (i) is immediate from Definition~\ref{m_rooted} and the observation that   
        $$Q^{-1}=I-E_{kn}=\begin{pmatrix}
        1 &  & & &  & 0 \\
         & 1 &  &      &  &  \\
         &  & \ddots & &  & -1 \\
         &  &        & &  &  \\
          &  & & & 1 &  \\
        0 &  & & &  & 1 \\
        \end{pmatrix},$$
    \end{proof}
 \end{comment}

    \begin{proof}
        and $Q^{-1}C'Q$ is
        $$\begin{pmatrix}
        c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
        \vdots \\
        c'_{k-11}     & c'_{k-1 2}           & \cdots     & c'_{k-1 n-1} & c'_{k-1k}+c'_{k-1n} \\
        c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{kn-1}-c'_{nk-1}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
        c'_{k+11}     & c'_{k+12}           & \cdots     & c'_{k+1\ n-1} & c'_{k+1k}+c'_{k+1n} \\
        \vdots              & \vdots & \ddots              & \vdots & \vdots \\
        c'_{n1}             & c'_{n2} & \cdots             & c'_{n\ n-1} & c'_{nk}+c'_{nn} \\
        \end{pmatrix}.$$
    \end{proof}

    {{Continue}}
    \begin{proof}
        (ii)
            By Lemma~\ref{v_rooted} $v'=Qu$ is $k$-rooted.  
            Since $Q^{-1}C'Qu=\rho(C')u$ by the assumption, we have
            $Q^{-1} C' Q u  = Q^{-1} \rho(C') Qu  =\rho(C')u$  \\
            $C'Qu=\rho(C')Qu$.

        (iii)
        Since $C'$ and $Q^{-1}C'Q$ have the same set of eigenvalues, clearly $\rho(C')$ = $\rho(Q^{-1}C'Q)$.

    \end{proof}

\begin{lem}\label{l_diag}   %end of label
If a square matrix $C'$ has a rooted eigenvector for $\lambda'$, then $C'+dI$ also has
the same rooted eigenvector for $\lambda'+d,$ where $d$ is a constant and $I$ is the identity matrix with the same size of $C'$.
\end{lem}

\begin{thm}
    Let $C$ be an $n\times n$ nonnegative Tmatrix. For $1\leq i \leq n$ and $1\leq j\leq n-1$, choose $c'_{ij}$
    such that $c'_{ij}\geq c_{ij}$ and $c'_{kj}\geq c'_{nj}>0$, and choose $r'_i$ such that $r'_i\geq c_{ik}+c_{in}$, and
    $r'_k \geq r'_n$. Moreover choose $c'_{in}:=r'_i-c'_{ik}$. Then $\rho(C)\leq \rho(C')$, when $C'=(c'_{ij})$.
\end{thm}

\begin{Proof}

The assumptions are necessary that $PCQ \leq PC'Q$, and C' is k-rooted, by (\ref{v_rooted}),
 For certain d, if C'+dI is k-rooted, then it has a k-rooted eigenvector with its spectral
  radius $\lambda + d$. C' would share the same eigenvector with C'+dI and has eigenvalue
   $\lambda$. So C'+dI and C+dI meet the conditions of (\ref{l_diag}), and we can show that
    $\rho(C' + dI) \geq \rho(C +dI)$ and then $\rho(C') \geq \rho(C)$  \qed    
\end{proof}

\subsection{Example}
    For the following $4\times 4$ matrix
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    we choose
    $$C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 &  0\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix}.$$
    Check the conditions $C[[n]|[n-1]]  \leq C'[[n]|[n-1]] $ \ref{thm_main},
    Then
    $\rho(C)\leq \rho(C')$ by previous theorem.

\subsection{Counterexample}
    For the following two $4\times 4$ matrices
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix},\quad C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 &  0\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$ 
    specify $n$=4, $k$=3 in $Q = I +E_{kn} = I + E_{34}$  
$$CQ=\begin{pmatrix}
0 & 0 & 1 & 2\\
1 & 0 & 0 & 1\\
1 & 1 & 0 & 0\\
1 & 1 & 1 & 1
\end{pmatrix},\quad C'Q=\begin{pmatrix}
0 & 0 & 1 & 2\\
1 & 0 & 1 & 1\\
1 & 1 & 0 & 0\\
1 & 1 & 1 & 1
\end{pmatrix},$$

    we have $CQ\leq C'Q$, but 
    $\rho(C)=2.234\not\leq 2.148= \rho(C')$. 
    This is because $c'_{33}+c'_{34}\not\geq c'_{43}+c'_{44}$.

\begin{thebibliography}{20}
\normalsize
\addcontentsline{toc}{chapter}{Bibliography}

\bibitem{chang}
Yen-Jen Cheng, {\it  A matrix realization of spectral bounds
of the spectral radius of a nonnegative matrix}, Ph.D. Thesis, NCTU, 2018.

\bibitem{prn_fros1}
A. E. Brouwer, W. H. Haemers, {\it Spectra of graphs}, Springer, 2012

\bibitem{prn_fros2}
R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.

\end{thebibliography}

\end{document}