### 2020-11-05 spectral radius of matrix new method/thesis/preliminiaries.tex

\documentclass[12pt]{report}%{article}

\topmargin 0pt

\topmargin=-1.5cm
\oddsidemargin=0.7cm
\textheight=23.5cm
\textwidth=15cm
\setlength{\baselineskip}{24pt}
\renewcommand{\baselinestretch}{1.5} %行距

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{comment}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
%\usepackage{colortbl}
\usepackage{enumitem}% enumerate labels   roman

\usetikzlibrary{arrows}
\setCJKmainfont{標楷體} %設定中文為系統上的字型，而英文不去更動，使用原TeX字型
\XeTeXlinebreaklocale "zh"             %這兩行一定要加，中文才能自動換行
\XeTeXlinebreakskip = 0pt plus1pt     %這兩行一定要加，中文才能自動換行
\title{Combinatorial Identities from Lagrange's Interpolation Polynomial}
\author{Student: Yen-Jung Huang  ~~~~~~~~~~~~~~~~~~~~~~~~~~Advisor: Chih-Wen Weng}
\date{} %不要日期

\def\UrlFont{\rm}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{remark}[thm]{Remark}
\newtheorem{eg}[thm]{Example}
\newtheorem{conj}[thm]{Conjecture}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{ex}[thm]{Exercise}
\newtheorem{prob}[thm]{Problem}
\newtheorem{exam}[thm]{Example}
\newtheorem{nota}[thm]{Notation}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ques}[thm]{Question}
\newtheorem{pof}[thm]{Proof.}

%\renewcommand {\refname} {Bibliography}

\begin{document}

\chapter{Preliminaries}

Let $\mathbb{R}$ and $\mathbb{C}$ denote the
 field of real numbers and complex numbers respectively.

\begin{defn}
    Let $C$ be an $n \times n$ real matrix, and $u \in \mathbb{R}^n$ be a
     nonzero column vector. The scalar $\lambda \in \mathbb{C}$ is an $\it{eigenvalue}$
      of C corresponding to the $\it{eigenvector}$ $u,$  if $Cu = \lambda u.$

\end{defn}

\begin{defn}

When $C$ is an $n \times n$ real matrix, the $\textit {spectral radius} $ $\rho(C)$ of $C$
 is defined by 
 $$\rho(C):=\max\{~|\lambda|~\ |~~\lambda\hbox{ is an eigenvalue of $C$}\},$$
where $|\lambda|$ is the magnitude of complex number $\lambda.$
\end{defn}

We are interested in spectral radius of the following matrix associated with a simple graph.

\begin{defn}
    Given an undirected graph G, the$\textit{ adjacency matrix}$ of G is the square
     matrix $A = (a_{ij})$ indexed by vertices of G, and
      \[a_{ij} =\begin{cases}
    1, \text{if $i$ is adjacent to $j$}, \\
    0, \text{otherwise.}
                \end{cases}
    \]
\end{defn}

\begin{defn}
Given an undirected graph G, the $\textit{spectral radius}$  $\rho(G) $ of G is the spectral
 radius of the adjacency matrix of G.
\end{defn}

    We introduce a notation of submatrix, which is taken from some columns and some rows of a
     matrix.

\begin{defn}
    For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column
    indices of $C$ respectively,  We use $C[\alpha|\beta]$ to denote the submatrix of $C$ with
    size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$
    and $j\in\beta$,
\end{defn}
The following theorem is a part of Perron-Frobenius theorem %\cite{'content'}.
\begin{comment}
$C[\alpha|\beta):=C[\alpha|\overline{\beta}],$ where $\overline{\beta}$ is the complement of $\beta$ in the set of column indices, and
similarly, for the definitions of $C(\alpha|\beta]$ and $C(\alpha|\beta).$
 For $\ell\in \mathbb{N},$ $[\ell]:=\{1, 2, \ldots, \ell\},$ symbol $-$ is the complete set of indices, and we use  $i$ to denote the singleton subset $\{i\}$ to reduce the double use of parentheses. For example of the $n\times n$ matrix $C$,
 $C[-|n)=C[[n]|[n-1]]$ is the $n\times (n-1)$ submatrix of $C$ obtained by deleting the last column of $C$.

    The bracket [] and parentheses () can be used together is the notation of sub-matrix, such as $C[k|e)$ or $C(k|e]$.
\end{comment}

\begin{thm}[Peron Frobius theorem]
        If $C$ is a nonnegative square matrix, then spectral radius $\rho(C)$ is an eigenvalue
         of $C$ with a corresponding nonnegative right eigenvector and a corresponding nonnegative
          left eigenvector.
\end{thm}

The following theorem is from \%. For completeness, we also provide the proof.  %\cite{chang}

\begin{thm}\label{pre_thm}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$, $P$ and $Q$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\Roman*)]
    \item \label{pre_thm_em1}  $PCQ\leq PC'Q$;   %
    \item \label{pre_thm_em2}  there exist a nonnegative column vector $u=(u_1, u_2, \ldots, u_n)^T$  and a
    scalar $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with
    associated eigenvector $Qu$;
    \item \label{pre_thm_em3}  there exist a nonnegative row vector $v^T=(v_1, v_2, \ldots, v_n)$  and a scalar
    $\lambda\in \mathbb{R}$such that $\lambda$ is an eigenvalue of $C$ with associated  left
    eigenvector $v^TP$; and
    \item \label{pre_thm_em4} $v^TPQu>0.$
\end{enumerate}
    Then $\lambda\leq \lambda'$. Moreover, $\lambda=\lambda'$if and only if
    \begin{equation}\label{pre0}
        (PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
    \end{equation}
\end{thm}

\begin{pof}
    Multiplying the nonnegative vector $u$ in theorem~\ref{pre_thm} assumption
    ~\ref{pre_thm_em1}, where $Qu$ is eigenvector of $C'$,  to the right of both terms of
    ~\ref{pre_thm_em1},
    \begin{equation}\label{pre1}
        PCQu\leq PC'Qu=\lambda'PQu.
    \end{equation}
    Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption
    ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), where $v^TP$ is
    left eigenvector of $C$ for $\lambda$, thus we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
    \end{equation}
        Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
         $\lambda\leq \lambda'$ and finish the proof of the first part.
    Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
    Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
    $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
    $u_j\not=0.$ Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=
    \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so $\lambda=\lambda'$ by (\ref{pre2}).

 \end{pof}

\end{document}