### 2020-12-04 spectral radius of matrix new method/thesis/introduction&front.tex
checked update 120420

\documentclass[12pt]{report}%{article}

\topmargin 0pt

\topmargin=-1.5cm
\oddsidemargin=0.7cm
\textheight=23.5cm
\textwidth=15cm
\setlength{\baselineskip}{24pt}
\renewcommand{\baselinestretch}{1.5} %行距

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
%\usepackage{colortbl}

\usetikzlibrary{arrows}
\setCJKmainfont{標楷體} %設定中文為系統上的字型，而英文不去更動，使用原TeX字型
\XeTeXlinebreaklocale "zh"             %這兩行一定要加，中文才能自動換行
\XeTeXlinebreakskip = 0pt plus1pt     %這兩行一定要加，中文才能自動換行
\title{Combinatorial Identities from Lagrange's Interpolation Polynomial}
\author{Student: Yen-Jung Huang  ~~~~~~~~~~~~~~~~~~~~~~~~~~Advisor: Chih-Wen Weng}
\date{} %不要日期

\def\UrlFont{\rm}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[chapter]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{remark}[thm]{Remark}
\newtheorem{eg}[thm]{Example}
\newtheorem{conj}[thm]{Conjecture}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{ex}[thm]{Exercise}
\newtheorem{prob}[thm]{Problem}
\newtheorem{exam}[thm]{Example}
\newtheorem{nota}[thm]{Notation}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ques}[thm]{Question}
\newtheorem{pof}[thm]{Proof.}

%\renewcommand {\refname} {Bibliography}

\begin{document}

%封面

\thispagestyle{empty}
\begin{center}
{ \Huge 國~~~~立~~~~交~~~~通~~~~大~~~~學}~\\~\\

\bigskip

{ \Huge 應~用~數~學~系}~\\~\\

\bigskip

{ \Huge 碩~~士~~論~~文}~\\~\\

\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip

{ \Huge 有向圖譜半徑之簡易估算方法}~\\~\\

\bigskip

{ \Huge A simple method on estimating spectral radius for some direted graphs}~\\~\\
~\\~\\~\\~\\~\\
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip \bigskip\bigskip\bigskip\bigskip\bigskip
\bigskip\bigskip\bigskip

{ \Large
\begin{tabular}{rcl}
研究生&：&陳科翰\\
指導教授&：&翁志文~教授
\end{tabular} }

\bigskip\bigskip
{ \Large 中~~華~~民~~國~~一~~百~~ㄧ~~十~~年~~一~~月 }
\large
\end{center}
\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\baselinestretch}{2} %行距
\thispagestyle{empty}
\begin{center}
{
\Large
有向圖譜半徑之簡易估算方法\\
A simple method on estimating spectral radius for some direted graphs\\~\\
\begin{tabular}{lccr}
Student: Ko-Han Chen  &&~~~& Advisor: Chih-Wen Weng\\
研究生:陳科翰　 &&~~~& 指導教授：翁志文~教授
\end{tabular}
}~\\

\bigskip

\renewcommand{\baselinestretch}{1} %行距

{ \LARGE 國~~~~立~~~~交~~~~通~~~~大~~~~學}\\~\\
{ \LARGE 應~用~數~學~系}\\~\\
{ \LARGE 碩~~士~~論~~文}\\~\\~\\~\\~\\
\renewcommand{\baselinestretch}{1} %行距
{ \large A Thesis

Submitted to Department of Applied Mathematics

College of Science

National Chiao Tung University

in Partial Fulfillment of Requirements

for the Degree of Master

in Applied Mathematics
\bigskip \medskip

January 2021

Hsinchu, Taiwan, Republic of China \bigskip \medskip

 中~~華~~民~~國~~一~~百~~ㄧ~~十~~年~~一~~月 }
\end{center}
\pagebreak

\label{abstract}

%中文摘要
\pagenumbering{roman}
\begin{center}
{  \LARGE
有向圖譜半徑之簡易估算方法
\bigskip\bigskip\bigskip

研究生：陳科翰  ~~~~~~~~~~ 指導教授：翁志文~教授 \\
國立交通大學  \\
\bigskip
應用數學系
\bigskip\bigskip\bigskip\bigskip
} \\~\\~\\~\\
\addcontentsline{toc}{chapter}{Abstract (in Chinese)}
{\large 摘~要}
\end{center}
 \bigskip

 找到一個簡單而且傑出的方法來估算譜半徑，試著縮小誤差，必定存在一個方法理性地證明它。我知道譜半徑是區分
 圖中點連線關係的重要指標，所以找一個簡單的方法來估算譜半徑是很有意義的。有了穩固的證明，我們可以估計連
 接矩陣的最小誤差譜半徑，並且找出存在這樣的上下界的充分條件。\\
\bigskip

\noindent 關鍵詞：譜半徑，簡易方法，有向圖，估算。
\pagebreak

%英文摘要
\begin{center}{\LARGE
A simple method on estimating spectral radius for some direted graphs
\bigskip\bigskip\bigskip}

{ \large
Student: Ko-Han Chen  ~~~~~ Advisor: Chih-Wen Weng \\
\Large

Department ~of~ Applied ~Mathematics
\bigskip

National~ Chiao ~Tung~ University
\bigskip\bigskip\bigskip\bigskip}\\
{\large Abstract}
\end{center}

%\begin{abstract}
\addcontentsline{toc}{chapter}{Abstract (in English)}
Find a simple and excellent executable method
 to estimate spectral radius, try to minimize the bios and there is a way to prove it sensible.
 I know that spectral radius is an important index to specify the relation of connected vertices in
  the graphs, so it is meanful to find a simple method to estimate spectral readius. With a solid proof,
   we can estimate the spectral radius of an adjacent matrix with minimal error and find out the sufficent
  condition on the existence of such bound.

\bigskip

\noindent {\bf Keywords}: spectral radius, simple method, directed graphs, estimation
%\end{abstract}
\pagebreak

\renewcommand{\baselinestretch}{1.2}
% 目錄
\large
\tableofcontents

\pagebreak

\label{Introduction}
% Introduction

\chapter{Introduction}
\normalsize
\pagenumbering{arabic}

\end{document}