### 2021-01-24 spectral radius of matrix new method/slides/oral0125.tex
before oral0125

\documentclass{beamer}

\usepackage{graphicx}
\usetheme{Berlin}
\usecolortheme{wolverine}

\graphicspath{{graphs/}}

%Information to be included in the title page:
\title{A simple method on comparison between spectral
radii of two directed graphs}
\author{陳科翰}
\institute{Mathematic department, NCTU}
\date{2021.01.25}

\begin{document}

\frame{\titlepage}

\begin{frame}
    \frametitle{References}
    \begin{thebibliography}{20}
        \normalsize
        \addcontentsline{toc}{section}{Bibliography}

        \bibitem{chang}
        Yen-Jen Cheng, {\it  A matrix realization of spectral bounds
        of the spectral radius of a nonnegative matrix}, Ph.D. Thesis, NCTU, 2018.

        \bibitem{spec_rad}
        A. E. Brouwer, W. H. Haemers, {\it Spectra of graphs}, Springer, 2012
        \bibitem{prn_fros2}
        R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.
        \bibitem{sage}
        [Sage] SageMath, the Sage Mathematics Software System (Version 9.2),
               The Sage Developers, 2020, https://www.sagemath.org.  %sage tutorial
        \bibitem{src_files}
        [github] K. H. Chen, C. W. Weng, source files to the essay, A simple method on comparison between spectral radii of two directed graphs,
             2021, https://reurl.cc/6yy32Z. % https://github.com/Urbaner3/NCTU_weng.

        \end{thebibliography}
\end{frame}

\begin{frame}
\frametitle{Introduction}
Let $\mathbb{R}$ and $\mathbb{C}$ denote the
 field of real numbers and complex numbers, respectively.
 Let $C$ be an $n\times n$ real square matrix. If there is
 a nonzero column vector $u\in\mathbb{C}^n$ such that
 $Cu=\lambda u$ for some scalar $\lambda\in\mathbb{C}$,
 then the scalar $\lambda$ is called the eigenvalue of
 $C$ with corresponding eigenvector $u$. And the spectral
 radius of a matrix $C$ is the largest magnitude (or complex
 modulus) of its eigenvalues, denoted by $\rho(C)$. We are
 interested in the spectral radius of the following matrix
associated with a simple directed graph.
\end{frame}

\begin{frame}
    \begin{definition}
        Given a directed graph $G$, the $\textit{adjacency matrix}$ of $G$ is the square
        matrix $A = (a_{ij})$ indexed by vertices of $G$, and
         \[a_{ij} =\begin{cases}
            1, \text{if $ji$ is an arc in $G$}, \\
            0, \text{otherwise.}
                \end{cases}
         \]
    \end{definition}
    Given a directed graph $G$, the spectral radius of $G$ is the
 spectral radius of the adjacency matrix of $G$, denoted by
 $\rho(G)$. Note that the spectral radius $\rho(G)$ is
 independent of the ordering of the vertex set of $G$.

\end{frame}

\begin{frame}
    In \cite{chang}, Cheng and Weng give many bounds of the spectral radius
     of a nonnegative square matrix. And based on their theory and Perron-Frobenius
      theorem, we give another approach to obtain an upper bound of the spectral
       radius and apply it on the adjacency matrix of a directed graph.

    All theorems come from continuous discussions between C.W. Weng and K.H. Chen.
     These were all documented.\cite{src_files}
\end{frame}

\begin{frame}
    \label{thm:Perron_Frobenius}
    \frametitle{Preliminaries}
    The following is Perron–Frobenius theorem, which provides a feature of
 nonnegative eigenvectors to nonnegative matrices.

\begin{theorem} \cite{prn_fros2} 
    If $C$ is a nonnegative square matrix, then the spectral radius $\rho(C)$ is an
    eigenvalue of $C$ with a corresponding nonnegative right eigenvector and a
    corresponding nonnegative left eigenvector.
    Moreover if $C$ is irreducible the above eigenvectors can be chosen to be positive. 
\end{theorem}
\end{frame}

\begin{frame}
    A well-known application of Theorem~\ref{thm:Perron_Frobenius}
 show that if matrix  $C'$ {\it majors} $C$ (in notation $C\leq C'$),
 i.e. $c_{ij}\leq c'_{ij}$ for all $ij,j$, then $\rho(C)\leq \rho(C')$.
Our main result shows that the assumption $C\leq C'$ can be a little loosen. 
Our theory is based on the following theorem, which is from \cite{chang}.
\end{frame}

\begin{frame}
    \begin{theorem}
        \label{pre_thm}
        Let $C=(c_{ij})$, $C'=(c'_{ij})$  be $n\times n$ real matrices
         with real eigenvalues $\lambda$, $\lambda'$ respectively such
         that there exist $n\times n$ matrices $P$ and $Q$   satisfying the following (i)-(iv).
    \begin{enumerate}[(i)]
        \item \label{pre_thm_em1}  $PCQ\leq PC'Q$;
        \item \label{pre_thm_em2} an eigenvector $Qu$ of $C'$  associated with
         eigenvalue $\lambda'$ exists for some nonnegative column vector
          $u=(u_1, u_2, \ldots, u_n)^T$.
        \item \label{pre_thm_em3} a left eigenvector $v^TP$ of $C'$ associated
         with eigenvalue $\lambda$ exists for some nonnegative row vector
         $v^T=(v_1, v_2, \ldots, v_n)$; and
        \item \label{pre_thm_em4} $v^TPQu>0.$
    \end{enumerate}
        Then $\lambda\leq \lambda'$. Moreover, $\lambda=\lambda'$ if and only if
        \begin{equation}\label{pre0}
            (PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
        \end{equation}
    \end{theorem}
\end{frame}

\begin{frame}
    Multiplying the nonnegative vector $u$ in assumption~\ref{pre_thm_em1}

    to the right of both terms of~\ref{pre_thm_em1},
    \begin{equation}\label{pre1}
       PCQu\leq PC'Qu=\lambda'PQu,
    \end{equation}
 where the above equality follows by $Qu$ being eigenvector of $C'$ for $\lambda'$.
\end{frame}

\begin{frame}
    Multiplying the nonnegative vector $v^T$ of $C$ in assumption
     ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), we have
    \begin{equation}\label{pre2}
        \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu,
    \end{equation}
    where the above first equality follows by $v^TP$ being
    left eigenvector of $C$ for $\lambda$.
    Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
        $\lambda\leq \lambda'$ and finish the proof of the first part.
\end{frame}

\begin{frame}
    Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
        Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
        $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
        $u_j\not=0.$ Conversely, (\ref{pre0}) implies
        \begin{align*} &\lambda v^TPCQu=v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j\\ =&
         \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu=\lambda'v^TPC'Qu,
         \end{align*} so $\lambda=\lambda'$ by (\ref{pre2}).
\end{frame}

\begin{frame}
    \frametitle{Our Method}
    We will apply Theorem~\ref{pre_thm} by using two particular square matrices $P$ and $Q$ to obtain our main result. 
We use $[n-1]$ as notation of the set of integers from $1$ to $n-1$, which is $\{1,2,\cdots,n-1\}$.
Throughout this thesis we fix $k\in [n-1]$, let $E_{kn}$ denote the $n\times n$ binary matrix with a unique $1$
appearing in the position $k,n$ of $E_{kn}$.
\end{frame}

\begin{frame}
    Now we apply the previous Theorem \ref{pre_thm} with $P=I$ and

    \begin{equation} \label{Q_1}
    Q=I+E_{kn}=\begin{pmatrix}
    1 &   &        &       &        &   & 0 \\
      & 1 &        &       &        &   &   \\
      &   & \ddots &       &        &   &   \\
      &   &        & 1     &        &   & 1 \\
      &   &        &       &\ddots  &   &  \\
      &   &        &       &        &1  & 0 \\
    0 &   &        &       &        &0  & 1
    \end{pmatrix},
    \end{equation}
\end{frame}

\begin{frame}
    so the matrix $PC'Q$ in assumption~\ref{pre_thm_em1} of Theorem~\ref{pre_thm} is
\begin{equation}\label{PC'Q}
  PC'Q=\begin{pmatrix}
         c'_{11} & c'_{12} & \cdots &  c'_{1k}+c'_{1n} \\
         c'_{21} & c'_{22} & \cdots &  c'_{2k}+c'_{2n} \\
         \vdots & \vdots & \ddots &  \vdots \\
         c'_{n1} & c'_{n2} & \cdots &  c'_{nk}+c'_{nn}
       \end{pmatrix},
\end{equation}
where $c'_{ij}$ denotes the $(i,j)$-entry of $C'$.
\end{frame}

\begin{frame}
    \begin{definition}%[$k$-rooted vector]
        A column vector $v'=(v'_1,v'_2,\ldots,v'_n)^T$ is called {\it $k$-rooted} if $v'_{j} \geq 0$ for $1 \leq  j \leq n$ and $v'_k\geq v'_n$.
        \end{definition}

        The following Lemma is immediate from the above definition.

        \begin{lemma}\label{lem:rt_vec}
        If $u=(u_1, u_2, \ldots, u_n)^T$, then
        \begin{enumerate}[(i)]
        \item \label{lem:rt_vec:en1} $Qu$ is $k$-rooted  if and only if $u$ is nonnegative;
        \item $u_k>0$ if and only if $(Qu)_k>(Qu)_n$.
        \end{enumerate}
        % \qed
        \end{lemma}
\end{frame}

\begin{frame}
    \begin{proof}
        (i), (ii) follow immediate from the definition of $k$-rooted and
         $Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$.
        \end{proof}

        Below is our first result, in which the first condition implies the first $n-1$ columns of $C'$ major
         the same columns of $C$, and the sum of $k$-th and $n$-th columns of $C'$ also majors that of $C$. The second and the third condition
         suggest that $C$ and $C'$ have nonnegative $k$-rooted eigenvectors. And the forth condition is simpler
          but with the same meaning in Theorem~\ref{pre_thm} 
\end{frame}

\begin{frame}
    We need a notation of submatrix, which is taken from some columns and some rows of a matrix.

\begin{definition}
    For a matrix $C=(c_{ij})$ and subsets $\alpha$, $\beta$ of row indices and column
    indices of $C$, respectively, we use $C[\alpha|\beta]$ to denote the
    submatrix of $C$ with size $ |\alpha| \times |\beta| $ that has entries $c_{ij}$ for $i\in \alpha$
    and $j\in\beta$. We use the notation $C[i|j]$ for short of $C[[i]|[j]].$
\end{definition}
\end{frame}

\begin{frame}
    \begin{theorem}\label{thm_main}
        Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ real matrices with real eigenvalues $\lambda$ and $\lambda'$ respectively.
    Assume that
        \begin{enumerate}[(i)]
        \item \label{thm_main:condition_i} $C[n|n-1]\leq C'[n|n-1]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
        \item \label{thm_main:condition_ii} there exists a $k$-rooted eigenvector vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ of $C'$ for $\lambda'$;
        \item \label{thm_main:condition_iii}there exists a nonnegative eigenvector vector $v^T=(v_1, v_2, \ldots, v_n)$ of $C$ for $\lambda$;
        \item \label{thm_main:condition_iv}$v^Tv'>0.$
        \end{enumerate}

        Then $\lambda\leq \lambda'$.
        Moreover, $\lambda=\lambda'$
        if and only if
        \begin{enumerate}[(a)]
            \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i>0$ and $v'_n>0;$
            \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1,$ $j \neq k$ with $v_i\ne 0 $ and $v'_j>0$;
            \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ with  $v_i>0$ and $ v'_{k}>v'_n$
        \end{enumerate} 
    \end{theorem}
\end{frame}

\begin{frame}
    \begin{proof}
        The proof is based on Theorem~\ref{pre_thm} with $P = I$ and $Q = I + E_{kn}$ in (\ref{Q_1}).
        The assumption \ref{pre_thm_em1} $PCQ\leq PC'Q$ of Theorem~\ref{pre_thm} holds by the condition \ref{thm_main:condition_i} of this theorem.
        Let $u = Q^{-1}v'$. Then $u$ is nonnegative and $C'Qu = \lambda' Qu$ by the condition \ref{thm_main:condition_ii} and \ref{lem:rt_vec:en1} in
         Lemma~\ref{lem:rt_vec}. Hence the assumption \ref{pre_thm_em2} of Theorem~\ref{pre_thm} holds. The assumptions \ref{pre_thm_em3} and \ref{pre_thm_em4}
          of Theorem~\ref{pre_thm} clearly hold by conditions~\ref{thm_main:condition_iii}, \ref{thm_main:condition_iv} of this theorem since $P = I$ and
           $v'= Qu$. Hence $\lambda \leq \lambda' $ by the necessary condition of Theorem~\ref{pre_thm}. Moreover,
            $\lambda = \lambda'$ if and only if (\ref{pre0}) holds, and this is equivalent to
             conditions \ref{thm_main:equ_cond_a}, \ref{thm_main:equ_cond_b} and \ref{thm_main:equ_cond_c} of this theorem.
    \end{proof}
\end{frame}

\begin{frame}
    We are interested in the matrices $C'$ that have $k$-rooted eigenvectors.
Motivated by the condition (i) of Theorem 2.3, we provide the following two definitions.
The first is the definition of $(k,n)$-sum.
\begin{definition}
    For an $n \times n$ matrix $C'=(c'_{ij})$, the $(k, n)$-{\it sum vector} of $C'$ is the vector
     obtained from the sum of the $k$-th and  $n$-th columns of $C'$.
\end{definition}
Note that the last column of $C'Q$ is the $(k, n)$-sum vector of $C'$.
Below is the definition of $k$-rooted matrix.
\begin{definition}\label{m_rooted}
    A  matrix $C'=(c'_{ij})$ is called {\it $k$-rooted}  if for
     each $i\not=k, n$ the  $i$-th column of $C'$  is $k$-rooted
     and the $(k,n)$-sum vector of $C'$ is $k$-rooted.
\end{definition}
\end{frame}

\begin{frame}
    We need a simple lemma for later use.

\begin{lemma}
$Q^{-1}=I-E_{kn}.$
\end{lemma}

\begin{proof}
Since $$Q(I-E_{kn}=(I+E_{kn})(I-E_{kn})=I-E_{kn}+E_{kn}-E_{kn}E_{kn}=I,$$ we have
$Q^{-1}=I-E_{kn}$.
\end{proof}

\end{frame}

\begin{frame}
    The matrix $Q^{-1}$ is explicitly written as
    $$Q^{-1}=I-E_{kn}=
\begin{pmatrix}
1 &   &        &       &        &   & 0 \\
  & 1 &        &       &        &   &   \\
  &   & \ddots &       &        &   &   \\
  &   &        & 1     &        &   & -1 \\
  &   &        &       &\ddots  &   &  \\
  &   &        &       &        &1  & 0 \\
0 &   &        &       &        &0  & 1
\end{pmatrix},$$

\end{frame}

\begin{frame}
    and if $C'=(c'_{ij})$ then $Q^{-1}C'Q$ has the form 
\begin{equation}\label{e6}
\begin{pmatrix}
            c'_{11}     & c'_{12} & \cdots     & c'_{1\ n-1} & c'_{1k}+c'_{1n} \\
            \vdots      & \vdots  & \vdots     & \vdots      & \vdots\\
            c'_{(k-1) 1}   & c'_{(k-1)  2}           & \cdots     & c'_{(k-1) (n-1)} & c'_{(k-1) k}+c'_{(k-1) n} \\
            c'_{k1}-c'_{n1} & c'_{k2}-c'_{n2} &\cdots      &c'_{k (n-1)}-c'_{n (k-1)}& c'_{kk}+c'_{kn}-c'_{nk}-x'_{nn}\\
            c'_{(k+1) 1}   & c'_{(k+1) 2}           & \cdots     & c'_{(k+1) (n-1)} & c'_{(k+1) k}+c'_{(k+1)
             n} \\
            \vdots              & \vdots & \ddots              & \vdots & \vdots \\
            c'_{n1}     & c'_{n2} & \cdots             & c'_{n (n-1)} & c'_{nk}+c'_{nn} \\
        \end{pmatrix}.\end{equation}

\end{frame}

\begin{frame}
    The following lemma shows that a $k$-rooted matrix has a $k$-rooted eigenvector.

\begin{lemma}\label{lma_m_rooted}
    Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(ii) hold.
        \begin{enumerate}[(i)]
            \item \label{lma_m_rooted_cond1} $C'$ is a $k$-rooted matrix if and only if $Q^{-1}(C'+dI)Q$ is nonnegative for some $d\geq 0$, where $I$ is the $n\times n$ identity matrix. 
            \item \label{lma_m_rooted_cond2} If $C'$ is $k$-rooted then there exists a  $k$-rooted eigenvector $v'$ of $C'$  for $\rho(C')$.
        \end{enumerate}
\end{lemma}
\end{frame}

\begin{frame}
    \begin{proof}[Proof.(i)]
        The matrix $Q^{-1}C'Q$ has $ij$ entry
        $$(Q^{-1}C'Q)_{ij}=\left\{
                             \begin{array}{ll}
                               c'_{ij}, & \hbox{if $i\not=k$ and $j\not=n$;} \\
                               c'_{kj}-c'_{nj}, & \hbox{if $i=k$ and $j\not=n$;} \\
                               c'_{ik}+c'_{in}, & \hbox{if $i\not=k$ and $j=n$;} \\ 
                               c'_{kk}+c'_{kn}-c'_{nk}-c'_{nn}, & \hbox{if $i=k$ and $j=n$},
                             \end{array}
                           \right.$$
        as shown in (\ref{e6}). 
        Hence $Q^{-1}(C'+dI)Q$ is nonnegative if and only if $C'$ is $k$-rooted by the definition of nonnegative matrix and $k$-rooted matrix.

    \end{proof}
\end{frame}

\begin{frame}
    \begin{proof}[Proof.(ii)]
        Suppose $C'$ is $k$-rooted. Choose $d\geq 0$ such that $Q^{-1}(C'+dI)Q$ is nonnegative.  
    Let $u$ be a nonnegative eigenvector of $Q^{-1}(C'+dI)Q=Q^{-1}C'Q+I$ for $\rho(C'+dI)=\rho(C')+d.$
    Note that $Q^{-1}C'Qu=\rho(C')u,$ and $Qu$ is $k$-rooted by Lemma~\ref{lem:rt_vec}. Hence $v'=Qu$ is what we want. 
    \end{proof}

\end{frame}

\begin{frame}
    Note that Theorem~\ref{thm_main} depends on eigenvectors. The following is an eigenvector-free Theorem.

\begin{theorem}\label{thm:conclusion}
    Let $C$ be an $n\times n$ nonnegative irreducible matrix and $C'$ be an $n\times n$ $k$-rooted matrix such that $C'Q$ majorizes $CQ$. 
 Then $\rho(C)\leq \rho(C')$. 
\end{theorem}

\begin{proof} Referring to (\ref{PC'Q}), the assumption \ref{thm_main:condition_i} in  Theorem~\ref{thm_main} holds. 
By Lemma~\ref{lma_m_rooted} \ref{lma_m_rooted_cond2}, there exists a  $k$-rooted eigenvector $v'$ of $C'$  for $\rho(C')$. 
Since $C$ is irreducible and nonnegative, there exists a positive eigenvector $v^T$ of $C$ for $\rho(C).$ 
Thus $v^Tv'>0$. Hence Theorem~\ref{thm_main} \ref{thm_main:condition_i}-\ref{thm_main:condition_iv} hold.
Hence  $\rho(C) \leq \rho(C')$ by   Theorem~\ref{thm_main}.
\end{proof}
\end{frame}

\begin{frame}
    \frametitle{Examples}
    Let $G$ be the digraph depicted below. 
    \begin{figure}
        \includegraphics[scale = 0.5]{exam3_10.PNG}
    \end{figure}

\end{frame}

\begin{frame}
    The following $4\times 4$ matrix is the adjacency matrix of $G$. 
    $$A=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix}.$$ 
We choose another matrix 
    $$A'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 & 0\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
which is the adjacency matrix of $G'$ depicted below. 
\end{frame}

\begin{frame}
    \begin{figure}
        \includegraphics[scale = 0.5]{exam3_11.PNG}
    \end{figure}

\end{frame}

\begin{frame}
    Note that both $G$ and $G'$ have the same number of edges, 
neither $A$ majories $A'$ nor $A'$ majories $A$. Our main result still can do comparison between 
$\rho(A)$ and $\rho(A')$. 
We first observe that $A$ is irreducible and $A'$ is $k$-rooted for $k=3$. Moreover 
$A'Q=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}\geq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}=AQ$. 
Hence Hence $\rho(A)\leq\rho(A')$ by Theorem~\ref{thm:conclusion} with $C=A$ and $C'=A'$. 
Both of the values of $\rho(C)$ and $\rho(C')$ are close to $2.511547$ by  calculating on computer. \cite[sage]{sage}
\end{frame}

\begin{frame}
    To finish the thesis, we provide an example to show that the `$k$-rooted' assumption of $C'$ is necessary in Theorem~\ref{thm:conclusion}.

    Consider the following two $4\times 4$ matrices
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix},\quad C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 &  0\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix}.$$
    With $n=4$ and $k=3$, we have  
    $$CQ=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 1
    \end{pmatrix}\leq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 1
    \end{pmatrix}=C'Q.$$
Using computer, \cite[sage]{sage}
    $$\rho(C)\approx 2.234\not\leq 2.148\approx \rho(C').$$
    This is because $C'$ is not $k$-rooted as $c'_{33}+c'_{34}=0\not\geq 1=c'_{43}+c'_{44}$.
\end{frame}

\end{document}